---
title: "{ecosystemom}"
author: "Bai Li <br><br> Contractor with ECS Federal LLC in support of <br> NOAA Fisheries Office of Science and Technology <br> Email: bai.li@noaa.gov"
institute: GOA-CLIM Monthly All-Hands Meeting
date: 2026/05/27
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css"]
    lib_dir: libs
    nature:
      ratio: '4:3'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "Page %current%"
---

layout: true

<style>
.progress-bar {
  position: fixed; left: 0; right: 0; bottom: 0;
  height: 8px; background-color: #0085CA; z-index: 100;
}

.left-wide {
  width: 65%;
  float: left;
}

.right-narrow {
  width: 30%;
  float: right;
}
</style>
  
```{r xaringanthemer, include=FALSE, warning=FALSE}
required_pkg <- c("xaringanthemer", "remotes", "webshot2", "xfun")
pkg_to_install <- required_pkg[!(required_pkg %in%
                                   installed.packages()[, "Package"])]
if (length(pkg_to_install)) install.packages(pkg_to_install)
lapply(required_pkg, library, character.only = TRUE)

remotes::install_github("nmfs-ost/nmfspalette")
library(nmfspalette)

style_xaringan(

  base_font_size = "15px",
  text_font_size = "1.5rem",

  title_slide_background_color = unname(nmfs_cols("darkblue")),
  title_slide_text_color = unname(nmfs_cols("white")),
  title_slide_background_size = "cover",
  title_slide_background_image = file.path("static", "slideswooshver.png"),

  background_image = file.path("static", "slideswoosh.PNG"),
  background_size = "cover",
  background_color = unname(nmfs_cols("white")),

  header_font_google = google_font("Arimo"),
  header_color = unname(nmfs_cols("darkblue")),

  text_color = unname(nmfs_cols("darkblue")),
  text_slide_number_color = unname(nmfs_cols("lightteal")),

  code_font_google = google_font("Source Code Pro"),
  code_highlight_color = unname(nmfs_cols("medteal")),

  inverse_background_color = unname(nmfs_cols("processblue")),
  inverse_text_color = unname(nmfs_cols("supltgray")),

  footnote_font_size = "0.6em",
  footnote_color = unname(nmfs_cols("darkblue")),
  footnote_position_bottom = "10px",

  link_color = unname(nmfs_cols("medteal")),


  extra_css = list(
    ".remark-slide-number" = list(
      "font-size" = "0.4em",
      "font-weight" = "bold",
      "margin" = "0px 840px -2px 0px"),

    ".title-slide h1, h2, h3" = list(
      "text-align" = "right"), 
    
    ".hyperlink-style" = list(
      "color" = "blue")
  )
)
```

.footnote[U.S. Department of Commerce | National Oceanic and Atmospheric Administration | National Marine Fisheries Service]

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

---
# Outline

- Fisheries Integrated Modeling System (FIMS) organization
- Why are we building {ecosystemom}?
- Features
- Main challenges
- Collaboration and contributions
- Ecosystem-to-stock-assessment simulation workflow demo

---
# Fisheries Integrated Modeling System (FIMS) Organization

```{r noaa-fims-architecture, echo=FALSE, out.width = "50%", crop=TRUE, fig.align='left'}
DiagrammeR::grViz("
digraph flowchart {
  node [shape = rectangle]
  A [label = 'NOAA-FIMS']
  B [label = 'FIMS']
  BB [label = 'FIMSRTMB']
  C [label = 'fishprior']
  D [label = 'ecosystemom', color='blue']
  E [label = 'FIMSdiags']
  A -> B
  A -> BB
  B -> BB
  A -> C
  A -> D
  D -> B
  A -> E
  B -> E
}
"
)
```

???
FIMS is a flexible suite of software tools to support sustainable fishery management

.footnote[
[1] FIMS website: .hyperlink-style[[https://noaa-fims.github.io/](https://noaa-fims.github.io/)]
]

---

# Why are We Building {ecosystemom}?

- 🏗️ Create a flexible ecosystem-to-stock-assessment simulation framework
- 🧪 Test FIMS using ecosystem models as operating models
- 🔍 Explore ecosystem-informed stock assessment approaches
- 🤝 Create interdisciplinary research and collaboration opportunities

---
# Features

Import and standardize outputs from ecosystem operating models

- Ecopath with Ecosim (EwE)
  - 🧪 Ecosim
  - 🛠️ Ecospace
  - 🔎 Individual-based version of Ecospace
- 🔎 Atlantis

---
# Example of Loading EwE outputs

```{r, echo=FALSE, fig.height=1, fig.width=1}
knitr::include_url("https://noaa-fims.github.io/ecosystemom/articles/ewe-ecosim-base-simulation.html#load-ewe-model-outputs")
```

---
# Features
Extract “true” population quantities from OMs, including annual or monthly trajectories and age compositions

---
# Example of Extracting Truth from EwE
```{r, echo=FALSE, fig.height=1, fig.width=1}
knitr::include_url("https://noaa-fims.github.io/ecosystemom/articles/ewe-ecosim-base-simulation.html#extract-truth-from-the-om")
```

---
# Features
Generate sampled observations from OM outputs for use in fisheries stock assessment models such as FIMS

- `sample_lognormal()`
- `sample_multinomial()`
- `sample_dirichlet_multinomial()`

---
# Example of Generating Sampled Data

```{r, echo=FALSE, fig.height=1, fig.width=1}
knitr::include_url("https://noaa-fims.github.io/ecosystemom/articles/ewe-ecosim-base-simulation.html#fishery-dependent-data")
```

---
# Features
Reshape environmental or diet composition data from OMs to support time series analyses such as Dynamic Structural Equation Models (DSEMs)

---
# Example of Creating Inputs for {dsem}
```{r, echo=FALSE, fig.height=1, fig.width=1}
knitr::include_url("https://noaa-fims.github.io/ecosystemom/articles/ewe-ecosim-base-simulation.html#create-dsem-inputs")
```

---
# Main Challenges
Different ecosystem models 

- produce different output formats
- has different spatial and temporal structures
- use different assumptions and parameterizations

---
# Collaboration and Contributions

.pull-left[
**💻Weekly coding sessions**

Mondays

- 3:30-5 PM ET
- 12:30-2 PM PT
- 9:30-11 AM HT

Starting June 1, 2026
]

.pull-right[
- Mix of development, training, and Q&A
- Topics provided beforehand 
- Work on {ecosystemom} and related workflows for sampling ecosystem model outputs to generate data for stock assessment models
]

---
# Ecosystem-to-Stock-Assessment Simulation Workflow Demo

Demo vignette: .hyperlink-style[[https://noaa-fims.github.io/ecosystemom/articles/ewe-ecosim-base-simulation.html](https://noaa-fims.github.io/ecosystemom/articles/ewe-ecosim-base-simulation.html)]


```{js, echo=FALSE}
(function() {
  var bar = document.createElement('div');
  bar.className = 'progress-bar';
  document.querySelector('.remark-slides-area').appendChild(bar);

  slideshow.on('afterShowSlide', function(slide) {
    var count = slideshow.getSlideCount();
    var index = slide.getSlideIndex();
    bar.style.width = (index / (count - 1) * 100) + '%';
  });
})();